---
title: "Time series"
author: "Vitor Borges Jr"
date: "19/07/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Basics - Time Series Modeling

There are few premisses we need to check before proceed with time series analyses.

## Stationary Series

To be calssified as stationary a series must meet three basic criterion:
        1. The mean of the series should not be a function of time rather should be a constant.
        2. The variance of the series should not be a function of time (homocedasticity).
        3. The covariance of the i th term and the (i + m) th term should not be a function of time.

## Why do I care about 'stationarity' of a time series?

## Random Walk



## 2. Explotration of time series data in R

Loading the data set

```{r}
data("AirPassengers")
```

The data serires must be in a time series format

```{r}
class(AirPassengers)
```

The strat of the time series

```{r}
start(AirPassengers)
```

The end of the time series

```{r}
end(AirPassengers)
```

The cycle of this time series is 12 months in a year

```{r}
frequency(AirPassengers)
summary(AirPassengers)
```

# Detailed metrics

plot of the time series - the number of passengers are distributed across the spectrum

```{r}
plot(AirPassengers)
```

Fit in a line

```{r}
plot(AirPassengers)
abline(lm(AirPassengers ~ time(AirPassengers)))
```

Print the cycle across years

```{r}
cycle(AirPassengers)
```

Aggregate the cylces and display a year on year trend

```{r}
plot(aggregate(AirPassengers, FUN = mean))
```

Box plot across months will give us a sense on seasona effect

```{r}
boxplot(AirPassengers ~ cycle(AirPassengers))
```

## Important inferences

1. The year on year trend cleaerly shows that the passengers have been increrasing without fail
2. The variance and the mean value in July and August is much higher than rest of the months
3. Even  though the mean value of each month is quite different their variance is small. Hence, we have strong seasonal effects with a cycle of 12 months or less.

## 3. Introduction to ARMA time series modeling

In ARMA, AR stands for autorregressive and MA stands for moving average

## Auto-regressive Time Series Model

## Moving Average Time Series Model

## Difference between AR and MA models

## Exploiting ACF and PACF plots

## 4. Framework and Application of ARIMA Time Series Modeling

## Overview of the framework

*How to do a time series analyses:*

1. Visualize the time series;
2. Stationarize the time series;
3. Plot ACF/PACF charts and find optimal parameters;
4. Build the ARIMA model;
5. Make predictions.

## Step 1: Visualize the time series

## Step 2 : Stationarize the series

## Step 3: Find optimal parameters

## Step 4: Build ARIMA model

## Step 5: Make predicitons

## Application of time series model

## Where did we start?

We need to remove unequal variances. We do this using log of the series. We need to address the trend component. We do this by taking difference of the series.

```{r, eval = TRUE, echo = FALSE}
adf.test(diff(log(AirPassengers)), alternative="stationary", k=0)

```

Next step is to find the right parameters to be used in the ARIMA model. We already know that the 'd' component is 1 as we need 1 difference to make the series stationary. We do this using the correlation plot.

```{r, eval = TRUE, echo = FALSE}
acf(diff(log(AirPassengers)))
pacf(diff(log(AirPassengers)))
```

Clearly, ACF plot cuts off after the first lag. Hence, we understood that value of p should be 0 as the ACF is the curve getting a cut off. While value of q should be 1 or 2. After a few iterations, we found that (0,1,1) as (p,d,q) comes out to be the combination with least AIC and BIC.

Let’s fit an ARIMA model and predict the future 10 years. Also, we will try fitting in a seasonal component in the ARIMA formulation. Then, we will visualize the prediction along with the training data. You can use the following code to do the same:

```{r, eval = TRUE, echo = FALSE}
(fit <- arima(log(AirPassengers), c(0, 1, 1),seasonal = list(order = c(0, 1, 1), period = 12)))

pred <- predict(fit, n.ahead = 10*12)

ts.plot(AirPassengers,2.718^pred$pred, log = "y", lty = c(1,3))
```


## 4. 